---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.


1. inspect the first few lines of the data set:

* what variables are there? of what type are the variables? what does each variable mean? what do we expect their data range to be?

The data is giving us the Parcel ID, Address, and style of the address given, the parcel ID is an char, the address is a char and style is a factor. There is also the occupancy which is a factor, sale date is a date, and price is a dbl
. There are many more variables in the data that are important to what is given and needed.


data range:
The data range can be the minimum price of the house and the maximum price. We 
can compare the highest house price and the lowest house price while focusing on the other variables that can affect the price. 

```{r}
library(classdata)
str(ames)
head(ames)

min_price <- min(ames$`Sale Price`)

max_price <- max(ames$`Sale Price`)


```


2. is there a variable of special interest or focus?

lot area is the variable of focus because we believe it will have the greatest influence on sale price. 

3. start the exploration with the main variable: sales price

* what is the range of the variable? draw a histogram for a numeric variable or a bar chart, if the variable is categorical. what is the general pattern? is there anything odd

* follow-up on oddities: see 4

```{r}

```

4. pick a variable that might be related to the main variable.
* sale price, occupancy, total living area, bedrooms, year built


* what is the range of that variable? plot. describe the pattern.

* what is the relationship to the main variable? plot a scatterplot, boxplot or facetted barcharts (dependening on the types of variables involved). Describe overall pattern, does this variable describe any oddities discovered in 3? Identify/follow-up on any oddities.


IDIVIDUAL VARIABLES: 
  Occupancy: 
Individually: follow through with your line of investigation. 
Include at least one plot that describes the relationship between sales prices and your variable. 
Include one paragraph describing the pattern you see. Are there oddities? 
Follow-up on (some of) them.
```{r}

```


  Total living area: 
Individually: follow through with your line of investigation. 
Include at least one plot that describes the relationship between sales prices and your variable. 
Include one paragraph describing the pattern you see. Are there oddities? 
Follow-up on (some of) them.
```{r}

```


  Bedrooms: 
Individually: follow through with your line of investigation. 
Include at least one plot that describes the relationship between sales prices and your variable. 
Include one paragraph describing the pattern you see. Are there oddities? 
Follow-up on (some of) them.
```{r}

```

  Year built: 
Individually: follow through with your line of investigation. 
Include at least one plot that describes the relationship between sales prices and your variable. 
Include one paragraph describing the pattern you see. Are there oddities? 
Follow-up on (some of) them.

```{r}

```


Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.
